Learning to Synthesize Rules for Network Intrusion Detectors

### 2020-01-11 paper/paper.tex
minor fixes to 4.3.1

\tname{} uses a composition of heurisitc functions to sort the rules
it produces. Each function uses a different criterion to estimate how
similar a rule created in the previous step is to existing Suricata
rules~\cite{emerging-threats-open}. We describe in the following each
of these heuristic functions. Each function $f_i$ ranges in the 0-1
domain. \tname{} uses the weighted average of these functions to sort
rules: $\sum{f_i}/|f_i|$.

\subsubsection{Size of a rule}
\tname{} uses the size of existing rules as a heuristic function. For
example, we noted that most rules in the dataset have 2 to 5 options,
so \tname{} values rules in this size range higher compared to other
sizes. Figure~\ref{fig:distribution-rule-size} shows the histogram of
rule sizes. The x-axis shows the number of options (\ie{}, size)
included in a rule and the y-axis shows the number of rules that have
that number of options. \tname{} uses this information to rank rules.

\begin{figure}[H]
  \centering
  \scalebox{1.2}{
    \begin{tikzpicture}
      \begin{axis}[
          ybar stacked,
          enlargelimits=0.09,
          legend style={at={(0.5,-0.15)},
            anchor=north,legend columns=-1},
          %          ylabel={\#rules},
          symbolic x coords={0, 1, 2, 3, 4, 5, 6, 7, 8, 9, 10},
          xtick=data,
          xticklabel style={rotate=45},          
          nodes near coords,
          every node near coord/.append style={font=\scriptsize},        
          nodes near coords align={vertical},
        ]
        \addplot coordinates {(0,11) (1,1301) (2,4910) (3,3662) (4,3250) (5,5353) (6,2911) (7, 1337) (8, 1246) (9,763) (10, 495)};
      \end{axis}
    \end{tikzpicture}
  }
  \vspace{-2ex}
  \caption{\label{fig:distribution-rule-size}Distribution of rule
    size.\Mar{how is it possible a rule with 0 options?}}
\end{figure}

\Mar{----------------------------------------- estou aqui}

Regras com um tamanho mais comum pontuam mais.
Nesse caso, a pontuação dada por esse componente será: \[\frac{rules\_per\_size[len(rule)]}{max(rules\_per\_size)}\]

>>>>

\tname{} uses a composition of heurisitc functions to sort the rules
it produces. Each function uses a different criterion to estimate how
similar a rule, created in the previous step, is to rules from a
public database~\cite{emerging-threats-open} including a total of
\numrulessuri\ rules. We describe in the following each of these
heuristic functions. Each function $f_i$ ranges in the 0-1
domain. \tname{} uses the weighted average of these functions to sort
rules: $\sum{f_i}/|f_i|$.

\subsubsection{Size of a rule}
One element that \tname{} uses to measure similarity of synthesized to
human-created rules is the size of the synthetic rule. Informally,
rules with popular sizes score high whereas inpopular rules score low.

Figure~\ref{fig:distribution-rule-size} shows the histogram of rule
sizes for our rule dataset. The x-axis shows the number of options
included in a rule (up to 10) while the y-axis shows the number of
rules with the corresponding number of options. Note that there are 11
rules with no options. These rules only use the header to signal a
traffic as malicious. Also note that most rules have 2 to 6
options. \tname{} values rules in that size range higher compared to
other sizes.

%\tname{} uses this information to rank rules.

\begin{figure}[H]
  \centering
  \scalebox{1.2}{
    \begin{tikzpicture}
      \begin{axis}[
          ybar stacked,
          enlargelimits=0.09,
          legend style={at={(0.5,-0.15)},
            anchor=north,legend columns=-1},
          %          ylabel={\#rules},
          symbolic x coords={0, 1, 2, 3, 4, 5, 6, 7, 8, 9, 10},
          xtick=data,
          xticklabel style={rotate=45},          
          nodes near coords,
          every node near coord/.append style={font=\scriptsize},        
          nodes near coords align={vertical},
        ]
        \addplot coordinates {(0,11) (1,1301) (2,4910) (3,3662) (4,3250) (5,5353) (6,2911) (7, 1337) (8, 1246) (9,763) (10, 495)};
      \end{axis}
    \end{tikzpicture}
  }
  \vspace{-2ex}
  \caption{\label{fig:distribution-rule-size}Distribution of rule
    size.}
\end{figure}

More precisely, we calculate the contribution of size of a given rule
$r$ with the following formula:

\[\frac{\mathit{rules\_per\_size}[\mathit{len(r)}]}{\mathit{max(rules\_per\_size})}\]

\noindent
, where $\mathit{len(r)}$ denotes the number of options of a rule $r$,
$\mathit{rules\_per\_size}$ encodes distribution of sizes as a
dictionary mapping sizes to number of associated rules, and
$\mathit{max(rules\_per\_size)}$ returns the maximum number of rules
for any dictionary entry. Consequently, the contribution of size is
proportional to its popularity. For example, the contribution of this
heuristic for a rule with 3 options would be 0.684 (=3,662/5,353)
whereas the contribution of a rule with 9 options would be 14.25
(=763/5,353). Note that the measurement is normalized in the 0-1
range.

\Mar{--------------------------- estou aqui}
